Python as a machine learning language: experiences from the scikit-learn project

### 2013-06-12 paper.tex
more an array processing: gensim comparison

\subsection{Data representation}

@@

The public API is oriented towards batch processing
rather than processing single samples per API call.
While classification and regression algorithms can indeed make predictions
for single samples,
\sklearn objects are not optimized for this use case.
(The few online learning algorithms implemented
are intended to take minibatches.)
Batch processing makes optimal use of NumPy and SciPy
by preventing the overhead inherent to Python function calls
and per-element dynamic type checking.
Although this might seem to be an artifact of the Python language,
and therefore an implementation detail that leaks into the API,
we argue that APIs should be designed
so as to not tie a library to a suboptimal implementation strategy,
and batch processing can enable fast implementations
in lower-level languages as well
(where locality of reference effects
and the possibility of internal parallelization come into play).
% Gilles: I don't think it is worth elaborating too much (if at all) on online learning. They are exceptions in scikit-learn.

@@

>>>>

\subsection{Data representation}
\label{sec:arrays}

@@

The public API is oriented towards batch processing
rather than processing single samples per API call.
While classification and regression algorithms can indeed make predictions
for single samples,
\sklearn objects are not optimized for this use case.
(The few online learning algorithms implemented
are intended to take minibatches.)
Batch processing makes optimal use of NumPy and SciPy
by preventing the overhead inherent to Python function calls
and per-element dynamic type checking.
Although this might seem to be an artifact of the Python language,
and therefore an implementation detail that leaks into the API,
we argue that APIs should be designed
so as to not tie a library to a suboptimal implementation strategy,
and batch processing can enable fast implementations
in lower-level languages as well
(where memory hierarchy effects
and the possibility of internal parallelization come into play).
% Gilles: I don't think it is worth elaborating too much (if at all) on online learning. They are exceptions in scikit-learn.

@@

Another package that deserves mention is the Gensim topic modeling toolkit
\citep{rehurek2010gensim}.
Although its stated goal is not machine learning,
it does exemplify a style of API design for scalable processing of ``big data''
that takes a different approach than the comparable models in \sklearn.
Gensim's way of dealing with large datasets is to use algorithms
that have $O(1)$ space complexity and can be updated online,
called \textit{document streaming} by \citeauthor{rehurek2010gensim}.
The API is designed around the Python concept of an \textit{iterable}
(supported in the language by a restricted form of co-routines called
\textit{generators}).
While the text vectorization part of \sklearn
also uses irerables to some extent,
the emphasis is still largely on (mini-)batch processing,
even in the truly stateless, O(1) memory vectorizers
that implement the hashing trick of \citet{weinberger2009}.
This way of processing, as argued earlier in Section~\ref{sec:arrays},
reduces various forms of overhead
and allows users to make use of the abundance of memory in modern machines.
No attempt is made to hide this batch-oriented processing from the user,
allowing control over the amount of memory actually dedicated
to \sklearn algorithms.
